Ionization of molecules by multicharged bared ions by using the stoichiometric model

### 2019-09-18 NewMonti_C.tex
changed colors

\documentclass[preprint,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}

\usepackage{authblk}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.50.0.2953}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{BibliographyScheme=Manual}
%TCIDATA{LastRevised=Thursday, March 07, 2019 14:11:39}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

%\input{tcilatex}
\begin{document}

\title{Ionization of molecules by multicharged bared ions by using the
stoichiometric model}
\author{author}
\affil{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio (CONICET-UBA).
Casilla de correo 67, sucursal 28 (C1428EGA) Buenos Aires, Argentina.}
%\author{author}
%\affil{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio (CONICET-UBA).
%Casilla de correo 67, sucursal 28 (C1428EGA) Buenos Aires, Argentina.}
%\affil{Fac. de Ciencias Exactas y Naturales, Universidad de Buenos Aires}
\date{\today }

@@

\section{Introduction}

The damage caused by the impact of multicharged heavy projectiles has 
become a field of interest due to its recent implementation in ion-beam 
cancer therapy. The effectiveness of the radiation depends on the 
choice of the ions. In particular, theoretical and experimental studies 
with different projectiles have concluded that charged carbon ions 
could be the most suitable ions to use. The study of such systems 
represent a challenge from the theoretical point of view. The most 
widely used method to compute ionization of multicharged atoms is the 
first Born approximation. 
This perturbative method warrants the $Z^{2}$ laws, where $Z$ is the 
projectile charge. However, the damage concentrated in the vicinities 
of the Bragg peak correspond to energies of hundreds of KeV/amu. 
Precisely in this region, often refered to as the intermediate energy 
region, the Born approximation starts to fail. Another theoretical 
issue arises from the targets themselves; we are dealing with complex 
molecules. This article deals precisely with these two aspects. 
First, we perform more appropriate calculations on the primary damage 
mechanism, i.e., atomic ionization by multicharged ions, which can 
replace the Born results. Second, we inspect and use an stoichiometric 
model, which reduces a molecule to a sum of atomic processes quantities 
weighted by the number of such atom in the molecule.

To overcome the first perturbative approximation limitations, and 
since the projectiles are multicharged ions, we resort to the 
Continuum Distorted Wave-Eikonal Initial State (CDW), which includes 
higher perturbative corrections. We start from the premise that the 
ionization process is the mechanism that deposits the most significant 
amount of primary energy. This process produces an electron-energy 
spectrum that needs to be integrated. Moreover, the ejected electrons 
become a new source of local damage. These secondary electrons are 
included in Monte Carlo simulations and hence their behavior must be
investigated. With this end in mind, we calculate not only the 
ionization cross-sections but also energy and mean angular distributions 
of the emitted electrons.

The molecular structure complexity of the target is dealt with by 
implementing the simplest stoichiometric model (SSM); we assume that 
the molecule is composed of isolated independent atoms. In all cases, 
the total cross-section is simply the sum of the cross-section of each 
atom.

Then, by implementing the SSM and the CDW --instead of the first Born 
approximation--, we calculate ionization cross-section of several 
molecules (see Table~\ref{tab:families}) by the impact of antiprotons,
H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$,\ and O$^{8+}$. Furthermore, 
in Section 3, we investigate different DNA and RNA molecules
such as adenine, cytosine, guanine, thymine, uracil, and also DNA 
backbone.

The results are processed to test the Toburen scaling rule, which 
states that the ratio between the ionization cross-section and the 
number of valence electrons (outer electrons) in terms of the 
projectile velocity can be arranged in a narrow universal band. 
We have also applied this rule to several hydrocarbons and nucleobases. 
In Section 3.2, we will prove that the width of the resulting universal 
band can be significantly reduced if we redefine the effective number 
of valence electrons.

It is well known that the residual electrons from the ionization 
process cause significant biological damage. To inspect this mechanism, 
in Section 3.3 and 3.4, we calculated the mean electron energy and
angular distributions. Surprisingly, we found a substantial dependence 
of the charged projectile, which is unexpected in the first Born 
approximation.

The hefty charge projectiles are dealt with the CDW. However, the 
stoichiometric model used seems to be very simplistic. The approximation
considers each atom as neutral, which is not correct. In Section 3.5, 
we used the molecular electronic structure code 
{\sc gamess}~\cite{gamess} to calculate the excess or defect of 
electron density one each atom. Then, the simple stoichiometric model is
modified to account for the departure from neutrality of the atoms. 
We find that for the DNA molecules this modification does not introduce 
a substantial change.

\section{Theory: Ionization of Atoms}

In our study, we will consider six atoms $\alpha=$ H, C, N, O, P, and S. 
Most of the organic molecules are composed of these atoms. Some 
particular molecules also include halogen atoms such as fluor and 
bromine; ionization cross-sections of these elements have been 
previously published~\cite{miraglia2008,miraglia2009}.

The total ionization cross-sections of these atoms were calculated using 
the CDW. The initial bound and final continuum radial wave functions 
were obtained by using the {\sc radialf} code, developed by Salvat and 
co-workers, and a Hartree-Fock potential obtained from the Depurated 
Inversion Model~\cite{mendez2016,mendez2018}. 
We used a few thousand pivot points to solve the Schr\"{o}dinger 
equation, depending on the number of oscillations of the continuum 
state. The radial integration was performed using the 
cubic spline technique. The number of angular momenta considered 
varied from 8, at very low ejected-electron energies, up to 30, 
for the highest energies considered. The same number of azimuth 
angles were required to obtain the four-fold differential 
cross-section. The calculation performed does not display prior-post 
discrepancies at all. Each atomic total cross-section in 
Eq.~(\ref{eq:sumion}) was calculated using 35 to 100 momentum transfer 
values, 28 fixed electron angles, and around 45 electron energies 
depending on the projectile impact energy. Further details of the 
calculation are given in Ref.~\cite{montanari2017}. Simultaneously, we 
will be reporting state to state ionization cross sections for these 
cases in Ref.~\cite{miraglia2019}, which will be very useful to estimate 
molecule fragmentation.

In Figure 1, we report our total ionization results for the six 
essential elements by the impact of six different projectiles: 
antiprotons, H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, and O$^{8+}$. 
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross-section scales with the square of the projectile 
charge, $Z^{2}$. The values of the impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 KeV. Moreover, we corroborated that the first Born 
approximation provides quite reliable results only for energies higher 
than a couple of MeV/amu. We use the same color to indicate the 
projectile charge throughout the figures of this work: dashed-red, 
solid-red, blue, magenta, olive and orange for antiprotons, 
H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, and O$^{8+}$, respectively. 
Notably, there is no complete tabulation of ionization of atoms by 
the impact of multicharged ions. We hope that the ones presented 
in this article will be of help for future works.

\section{Ionization of Molecules}

\subsection{The stoichiometric model}

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.
\begin{table}
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.65\textwidth}|}
\hline
 CH  & CH$_4$, C$_2$H$_2$, C$_2$H$_4$, C$_2$H$_6$, and C$_6$H$_6$ \\
\hline
 CHN & C$_5$H$_5$N, C$_4$H$_4$N$_2$, C$_2$H$_7$N, and CH$_5$N \\
\hline
 \multirow{4}{*}{DNA} & adenine (C$_5$H$_5$N$_5$), 
                        cytosine (C$_4$H$_5$N$_2$O), \\
     & guanine (C$_5$H$_5$N$_5$O), thymine (C$_5$H$_6$N$_2$O$_2$), \\
     & uracil (C$_4$H$_4$N$_2$O$_2$), DNA backbone (C$_5$H$_{10}$O$_5$P), \\
     & dry DNA (C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$)\\
\hline
\end{tabular}
\caption{Sixteen molecular targets of our interest classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

In Figure 2, we report the total ionization cross sections by the 
impact of multicharged ions for the four fundamental DNA components: 
adenine, cytosine, guanine and thymine. Our reports for uracil and 
DNA backbone containing P are given in Figure 3. For adenine, the 
agreement with the experimental data available in our range of validity 
is quite good. For uracil, we have a puzzling situation: our results 
show good agreement with the experiments by proton impact of Itoh 
{\it et al.}~\cite{itoh2013}, but for the same target our theory fails 
by a factor of two for the impact of C$^{4+}$ and O$^{6+}$ ions. 
Nonetheless, it should be stated that our theoretical results coincide 
with calculations by Champion, Rivarola and 
collaborators~\cite{champion2012,agnihotri2012}, which may indicate a 
possible misstep of the experiments.

@@

The first attempt to develop a comprehensive but straightforward 
phenomenological model for electron ejection from large molecules was 
proposed by Toburen {\it et al.}~\cite{toburen1975,toburen1976}. 
The authors found it convenient to scale the experimental ionization 
cross section in terms of the number of outer or weakly bound valence 
electrons (total number of electrons minus the inner-shell ones). 
We introduce the cross section per weakly bound electron or ionization 
cross section per valence electron, $\sigma_{e}$, as
\begin{equation}
\sigma_{e}=\frac{\sigma_{M}}{N_{M}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}}{\sum\limits_{\alpha}n_{\alpha}\nu_{\alpha}}
=\sigma_{e}(v)\,, 
\label{27} 
\end{equation}
where
\begin{equation}
\nu_{\alpha}=\left\{ 
\begin{array}{ll}
4, & \text{for C,} \\ 
5, & \text{for N and P,} \\ 
6, & \text{for O and S,}
\end{array}\right.
\label{eq:nelec} 
\end{equation}
and $\nu_{\text{H}}=1$. The Toburen rule can be stated by saying that 
$\sigma_{e}$ is a \textit{universal} parameter independent on the 
molecule, which depends solely on the impact velocity. 
In Ref.~\cite{toburen1976}, it was found experimentally that for proton 
impact on some simple molecules, $\sigma_{e}$ results
\begin{equation}
\begin{array}{cccc}
E\text{(MeV)}               & 0.25 & 1.00 & 2.00 \\ 
\sigma_{e} (10^{-16}cm^{2}) & 0.39 & 0.17 & 0.11
\end{array}
\label{30}
\end{equation}
with an estimated error of about 20\%. A similar ratio was found in 
Ref.~\cite{itoh2013} for proton impact on uracil and adenine. 
Figure 4a shows $\sigma_{e}$ calculated with the CDW as a function of 
the impact velocity for different projectile charges computed with the 
SSM for the sixteen molecular targets displayed in 
Table~\ref{tab:families}. The universality with $Z$ is the one provided 
by Born approximation, i.e., $\sigma _{e}(Z)=Z^{2}\sigma_{e}(Z=1)$, 
and it holds for large impact velocities, as shown Figure 4a.
Of course, for lower impact velocities, the CDW breaks the behavior of 
the $Z^{2}$ rule. Although the Toburen rule holds for high energies, 
its performance is still not satisfactory: the universal band is quite 
broad. The departure of our theoretical 
results from the Toburen rule can be easily explained by the fact that 
our atomic cross sections $\sigma_{\alpha}$ behave differently for 
each atom in terms of the projectile charges. By inspecting Figure 1, 
one can easily see that the rule $\sigma_{\alpha}/\nu_{\alpha}\sim
\sigma_{e}$ is not well satisfied by the CDW. A much better general 
atomic rule is given by
\begin{equation}
\sigma_{e}^{\text{CDW}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}^{\text{CDW}}}{\sum\limits_{\alpha}n_{\alpha}
\nu_{\alpha}^{\text{CDW}}},
\label{32} 
\end{equation}
where 
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
4,\ \ \  & \text{for C, N, and O,} \\ 
4.5,\ \  & \text{for P and S}
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}
and, obviously, $\nu_{\text{H}}^{\text{CDW}}=1$. The cross sections 
$\sigma_{e}^{\text{CDW}}$ are plotted in Figure~4b. A much better sharp 
band is observed, especially at high energy, where the theory is 
expected to work. It will be interesting for experiments to cross-check 
this theoretical prediction.

\subsection{Emitted electron energies}

In a given biological medium, direct ionization by ion impact accounts 
for just a fraction of the overall damage. Secondary electrons, as well 
as recoil target ions, also contribute substantially to the total damage. 
We can consider the single differential cross section of the shell 
$nl$ of the atom $\alpha$, $d\sigma _{\alpha nl}/dE$, to be a function 
of the ejected electron energy $E$ as a simple distribution 
function~\cite{surdutovic2018}. Then, we can define the mean value 
$\overline{E}_{\alpha}$ as in Ref.~\cite{abril2015},
\begin{eqnarray}
\overline{E}_{\alpha} &=&\frac{\langle E_{\alpha}\rangle}{\langle
1\rangle}=\frac{1}{\sigma_{\alpha}}\sum\limits_{nl}\int dE\,E
\frac{d\sigma_{\alpha,nl}}{dE}\,,  
\label{40} \\
\langle 1\rangle &=&\sigma_{\alpha}=\sum\limits_{nl}\int dE\,
\frac{d\sigma_{\alpha,nl}}{dE}\,,  
\label{50}
\end{eqnarray}
where $\Sigma_{nl}$ takes into account the sum of the different 
sub-shell contributions of the element $\alpha$.

Figure 5 shows $\overline{E}_{\alpha}$ for six elements from 
Table~\ref{tab:families}. The range of impact velocities was shorten up 
to $v=10$ a.u. due to numerical limitations in the spherical harmonics 
expansion. In our theoretical treatment, we expand our final continuum 
wave function as per usual,
\begin{equation}
\psi_{\overrightarrow{k}}^{-}(\overrightarrow{r})=\sum_{l=0}^{l_{\max
}}\sum_{m=-l}^{l}R_{kl}^{-}(r)Y_{l}^{m}(\widehat{r})Y_{l}^{m^{\ast }}
(\widehat{k})\,.
\label{60}
\end{equation}
We are confident with our calculations up to $l_{\max}\sim 30$. 
As the impact velocity $v$ increases, so do $\langle E_{\alpha}\rangle$
and $l_{\max}$, which results in the inclusion of very oscillatory 
functions in the integrand. Furthermore, the integrand of
$\langle E_{\alpha}\rangle$ includes the kinetic energy $E$
(see Eq.~(\ref{40})), which cancels the small energy region and 
reinforces the large values, making the result more sensible to large
angular momenta. Regardless, for $v>10$ a.u., the first Born 
approximation holds.

In Figure 5, we estimate $\overline{E}_{\alpha}$ in the 0.5--2 a.u.
velocity range, or equivalently from 15 to 50 eV, for all the targets.
Our results agree with the experimental findings~\cite{surdutovic2018}. 
The dependence of the mean energy value with the projectile charge $Z$ 
is surprisingly sensible, which can duplicate the proton results. 
This effect can be attributed to the depletion caused by the 
multicharged ions to the yields of low energy electrons. In the high 
electron energy regime, the CDW falls on the simple first Born 
approximation, surviving the $Z^{2}$ law. Then, the ratio in 
Eq.~(\ref{40}) cancels out and $\overline{E}_{\alpha}$ becomes a 
universal value independent on $Z$.

Extending the simple stoichiometric model for the mean electron energy
calculation, it results
\begin{equation}
\overline{E}_{M}=\frac{\sum\limits_{\alpha}n_{\alpha}
\overline{E_{\alpha}}}{\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}}.
\label{70}
\end{equation}
For impact of H$^{+}$ and He$^{2+~}$ on water at 1 MeV/amu, we obtain 
$\overline{E}_{\text{H}_{2}\text{O}}=43.7$ and 45 eV, while the 
experimental values 
on liquid water by Pimblott and LaVerne [xxx] were found to be 51.5 and 
52.2 eV, respectively. Just to stress the importance of the projectile 
charge, for O$^{8+}$ on water, at the same impact energy 1 MeV/amu, 
we obtained $\overline{E}_{\text{H}_{2}\text{O}}=54.7$ eV, which is 25\% 
larger than 
proton impact result. It is worth noting that there should not be any 
difference if the calculations were carried in first Born approximation. 
We will came back on this issue later.

\subsection{Emitted electron angles}

As mentioned before, secondary electrons contribute to the total damage. 
Then, not only is the ejection energy important but also the angle 
of emission. Once again, we can consider the single differential cross 
section in terms of the ejected electron solid angle $\Omega$, 
$d\sigma_{\alpha,nl}/d\Omega$, to be expressed as a distribution function, 
and the mean angle $\overline{\theta}_{\alpha}$ can be defined as
\begin{equation}
\overline{\theta}_{\alpha}=\frac{\langle\theta_{\alpha}\rangle}
{\langle 1\rangle}=\frac{1}{\sigma_{\alpha}}\sum\limits_{nl}
\int d\Omega\,\theta\,\frac{d\sigma_{\alpha,nl}}{d\Omega}
\end{equation}
Figure 6 shows $\overline{\theta}_{\alpha}$ for our six elements of 
interest. A new important dependence of $\overline{\theta}_{\alpha}$ 
with $Z$ is observed. It is a general belief~\cite{Rudd1992} that the 
angular dispersion of emitted electrons are nearly isotropic. This effect 
is caused by the insignificant angular anisotropy of sub-50-eV yield. 
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. But, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region, as observed in Figure 6. 
For example, for C$^{6+}$ impact, the Bragg peak occurs at 0.3 MeV/amu, 
where $\overline{\theta}_{\alpha}$, computed with the CDW method, is 
about half of the value obtained with the first Born approximation. 
This correction should close the damage to the forward direction.

We can attribute this correction to the capture to the continuum effect;
the larger the charge $Z$, the smaller $\overline{\theta}$ will be. Of 
course, this effect only holds at intermediate energies and not at high 
impact energies, where the Born approximation rules. One illustrative 
observation is the behavior of antiprotons: the projectile in this case 
repels the electrons, making the distribution almost symmetric. 
Note the opposite effect of proton and antiprotons; they run one 
opposite to the other, as compared with the first Born approximation.

\subsection{A modified stoichiometric model}

The SSM considers the molecule to be assembled by isolated neutral atoms, 
which is definitively unrealistic. A first improvement can be suggested 
by assuming that the atoms are not neutral and that they have an uneven
distribution of electrons within the molecule; this distribution can be
given by an effective charge $q_{\alpha}$. A possible value for 
$q_{\alpha}$ is given by the Mulliken charge. However, there are a wide
variety of charge distributions, such as the net or natural atomic
charge~\cite{lee2003}, the L\"owdin charge, etc.

Consider that the total amount of electrons $Q_{\alpha }$ on the element
$\alpha$ are equally distributed on all the $\alpha$ atoms. Therefore, 
each element $\alpha$ will have an additional charge: 
$q_{\alpha}=Q_{\alpha}/n_{\alpha}$, which can be positive or negative.
This value will depend on the relative electronegativity value with 
respect to the other atoms~\cite{rappe1991}. Now, instead of an
integer number of elements $n_{\alpha}$ of the atom $\alpha$, we have a 
fractional number of atoms given by 
\begin{equation}
n_{\alpha }^{\prime }=n_{\alpha }-
\frac{q_{\alpha }}{\nu_{\alpha }^{\text{CDW}}}
\label{eq:newstoi}
\end{equation}%
In the case of neutral atoms, $q_{\alpha}=0$ and we recover 
$n_{\alpha}^{\prime}=n_{\alpha}$, as it should be.

To inspect the effect of $q_{\alpha}$, we computed the molecular
structure of several nucleobases with the {\sc gamess} code. We 
used the 6-31G(d,p) basis set, which includes polarization 
functions for all the atoms. The calculations were carried out 
implementing the B3LYP functional~\cite{Becke1993,Stephens1994} to 
account for the correlation and exchange effects. 
In Table~\ref{tab:newstoi}, we display the charge $q_{\alpha}$ of four 
DNA molecules. The last column shows the new stoichiometric formula
found by implementing Eq.~(\ref{eq:newstoi}).
\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.12\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}|p{
0.07\textwidth}|p{0.07\textwidth}|p{0.25\textwidth}|}
\hline
Element & C & H & N & O & New stoichiometry \\
\hline
Adenine & +0.32 & +0.23 & --0.55 &       & C$_{4.92}$H$_{4.77}$N$_{5.14}$ \\ 
\hline
Cytosine & +0.28 & +0.21 & --0.56 & --0.53 & C$_{3.93}$H$_{2.79}$N$_{5.14}$O$_{1.13}$ \\ 
\hline
Guanine & +0.46 & +0.20 & --0.58 & --0.36 & C$_{4.89}$H$_{4.80}$N$_{5.15}$O$_{1.09}$ \\ 
\hline
Thymine & +0.20 & +0.19 & --0.54 & --0.52 & C$_{4.95}$H$_{1.95}$N$_{6.13}$O$_{2.13}$ \\ 
\hline
\end{tabular}
\caption{Effective charge $q_{\alpha}$ and new stoichiometric formula 
defined by Eq.~(\ref{eq:newstoi}) for four DNA molecules.}
\label{tab:newstoi}
\end{center}
\end{table}

The previous formula can be rewritten by replacing the integer value 
$n_{\alpha}$ by the fractional values $n_{\alpha}^{\prime}$. This 
analog formulations are now implemented to compute the new ionization 
cross sections, $\sigma^{\prime}$.

In Figure 8, we display the relative difference of ionization cross 
sections 
\begin{equation}
e=\frac{\sigma^{\prime}-\sigma}{\sigma}\times 100
\label{120}
\end{equation}
for the four DNA bases. The difference obtained is of just very few 
percents, which indicates that the SSM is a quite robust model to treat 
the type of molecule we are dealing with, and of course within the 
range error that one can expect from the SSM.

Our molecular binding energies of the outer electrons for adenine, 
cytosine, guanine and thymine are shown in Figure~7. On the left side 
of the figure, we show the atomic Hartree--Fock energies of the 
constituent elements, which can give an idea of how the molecular levels 
distribute. A dashed line around -0.8 a.u. is drawn to separate in two 
the band levels of the outer electrons. The electrons in the molecular 
targets corresponding to the $2s$ of N and C in the atomic case are 
placed in a secondary gap. The lower levels of the secondary gap 
correspond to the oxygen. The number of electrons in the valence 
shell for all O would seem to be equal to 4, as previously setted by the 
scaling rule Eq.~(\ref{eq:scalingCDW}). The N case is not as straightfoward;
the electrons that would correspond to the $2s$ are located just below
the line. The $\nu_{\alpha }^{\text{CDW}}$ value given for N would 
suggest that a significant amount of electron density from the secondary gap 
is shared with the valence band.

\begin{figure}[H]
\centering
\includegraphics[width=0.95\textwidth]{energylevels/DNA.eps}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We now return to the calculation of the mean electron energy of H$^{+}$ 
and He$^{2+}$ in water, $\overline{E}_{\text{H}_2\text{O}}$. If we 
consider that the residual charge of each hydrogen in water is 
$q_{\text{H}}=+0.35$~\cite{rappe1991}, which implies that on oxygen we have a 
residual charge of $q_{\text{O}}=-0.7$, we could re-write the formula of water 
using Eq.~(\ref{120}) as H$_{1.3}$O$_{1.17}$ by considering 
$n_{\text{O}}^{\text{CDW}}\simeq 4$ and 
$n_{\text{H}}^{\text{CDW}}\simeq 1$. Recalculating mean electron energy 
by impact of H$^{+}$ and He$^{2+}$ on water at 1 MeV/amu, we obtain
$\overline{E}_{\text{H}_2\text{O}}=46.2$ and 47.6 eV. Our results 
are 10\% apart from the experimental values: 51.5 and 
52.2~eV~\cite{pimblott2007}, respectively, which is not bad considering 
the simplicity of our model.

Another parameter that can computed within our model is the fraction of
energy carried by the electron $f$, which is defined as
\begin{equation}
f=\frac{\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}
\overline{E_{\alpha}}}{\sum\limits_{\alpha}n_{\alpha}(\sigma_{\alpha}
\overline{E_{\alpha}}+\sum\limits_{nl}\sigma_{\alpha nl}E_{\alpha nl})}\,.
\label{90}
\end{equation}
Using our shell-to-shell ionization cross sections 
$\sigma_{\alpha nl}$~\cite{miraglia2019}, and atomic binding energies of 
the atom $\alpha$, $E_{\alpha nl}$~\cite{clementi}, we obtain 
$f=0.70$ and $0.71$ on water at 1 MeV/amu for H$^{+}$ and He$^{2+}$, 
which differ from the respective experimental values $f=0.81$ and $0.80$.

The difference between our calculations and the experiments could be 
attributed to the binding energy values of the molecular water~12.65 eV 
and its atomic components, H (13.6 eV) and O (17.19 eV). This theoretical
discrepancy goes beyond our independent atom approximation and so it 
cannot be accounted for by our model. Once again, we could estimate this 
contribution by resorting to the stoichiometric model. We can make an 
asses for a maximum difference per molecule of water as
\begin{equation}
\delta E=(13.6-12.65)\times1.17+(17.19-12.65)\times1.3=7\,\text{eV}  
\label{80}
\end{equation}
Most of this energy should be transfer to the kinetic energy of the 
emitted electron. Thus, by adding 7 eV to the previous values, we obtain
$\overline{E}_{\text{H}_2\text{O}}=53.2$ and 54.6 eV for H$^{+}$ and 
He$^{2+}$ on water at 1 MeV/amu, respectively, which are much closer to
the experimental values. If we add $\delta E$ to the numerator of 
Eq.~(\ref{90}), we obtain $f=0.81$ and $0.80$, in perfect agreement with 
the experiments (see Table 1 of Ref.~\cite{pimblott2007}).

>>>>

\documentclass[preprint,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{gensymb}
\usepackage{float}
%\usepackage{authblk}
%\setcounter{MaxMatrixCols}{10}

\begin{document}

\title{Ionization of molecules by multicharged bared ions by using the
stoichiometric model}
%\author{author}
%\affiliation{Consejo Nacional de Investigaciones Cient\'{\i}ficas y
%T\'{e}cnicas - Universidad de Buenos Aires, Instituto de
%Astronom\'{\i}a y F\'{\i}sica del Espacio, Pabell\'on IAFE, 1428 Buenos Aires, Argentina;}

\date{\today }

@@

@@

\subsubsection{Toburen numbers}

The first attempt to develop a comprehensive but straightforward
phenomenological model for electron ejection from large molecules was
proposed by Toburen {\it et al.}~\cite{toburen1975,toburen1976}.
The authors found it convenient to scale the experimental ionization
cross section in terms of the number of outer or weakly bound valence
electrons \textbf{(i.e. total number of electrons minus the K-shell)}.
\textbf{Following Toburen we can define the ionization} cross section per weakly bound electron
%or ionization cross section per valence electron,
$\sigma_{e}^T$, as
\begin{equation}
\sigma_{e}^T=\frac{\sigma_{M}}{N_{M}}=\frac{\sum\limits_{\alpha}
n_{\alpha} \sigma_{\alpha}}{\sum\limits_{\alpha}n_{\alpha}\nu_{\alpha}^T}
=\sigma_{e}(v)\,,
\label{27}
\end{equation}
where \textbf{$\nu_{\alpha}^T$ are the Toburen numbers for each atom:}
\begin{equation}
\nu_{\alpha}^T=\left\{
\begin{array}{ll}
1, & \text{H,} \\
4, & \text{for C,} \\
5, & \text{for N and P,} \\
6, & \text{for O and S,}
\end{array}\right.
\label{eq:nelec}
\end{equation}
%and $\nu_{\text{H}}=1$.
The Toburen rule can be stated by saying that
$\sigma_{e}$ is a \textit{universal} parameter independent on the
molecule, which depends solely on the impact velocity\textbf{
, and holds for high impact energies (i.e. 0.25-5 Mev/amu).}
%In Ref.~\cite{toburen1976}, it was found experimentally that for proton
%impact on some simple molecules, $\sigma_{e}$ results
%\begin{equation}
%\begin{array}{cccc}
%E\text{(MeV)}               & 0.25 & 1.00 & 2.00 \\
%\sigma_{e} (10^{-16}cm^{2}) & 0.39 & 0.17 & 0.11
%\end{array}
%\label{30}
%\end{equation}
%with an estimated error of about 20\%.

\textbf{
These $\nu_{\alpha}^T$ can be interpreted as the number of active 
electrons in the collision. Of course, at very high energies also de 
K--shell electrons will be ionized and these numbers will be different. }

A similar \textbf{dependence with the number of weakly bound electrons} was found in
Ref.~\cite{itoh2013} for proton impact on uracil and adenine.

Figure 4a shows \textbf{our CDW results for} $\sigma_{e}^T$ as a function of
the impact velocity for different projectile charges computed with the
SSM for the sixteen molecular targets displayed in
Table~\ref{tab:families}. The universality with $Z$ is the one provided
by Born approximation, i.e., $\sigma _{e}^T(Z)=Z^{2}\sigma_{e}^T(Z=1)$,
and it holds for large impact velocities, as shown Figure 4a.
Of course, for lower impact velocities, the CDW breaks the behavior of
the $Z^{2}$ rule.

Although the Toburen \textbf{scaling} holds for high energies,
its performance is still not satisfactory: the universal band is quite
broad.

\subsubsection{New scaling}

The departure of our theoretical
results from the Toburen rule can be explained by %the fact that
%our atomic cross sections $\sigma_{\alpha}$ behave differently for
%each atom in terms of the projectile charges. By
\textbf{inspecting Figure 1. It can be noted}
 that the rule $\sigma_{\alpha}/\nu_{\alpha}^T\sim
\sigma_{e}^T$ \textbf{approximatelly constant} is not well satisfied by 
the CDW model.
\textbf{
For example, Figure 1 shows that the cross sections for O are actually 
very similar to the cross sections for C, suggesting 4 active electrons 
in O instead of 6. In the same way the number of active electrons for
N, P and S are also different from the $\nu_{alpha}^T$ in Equation 
(\ref{eq:nelec}). This will be an important change in the scaling for 
molecular targets. Based on the CDW results we propose a new scaling 
given by}
%A much better general atomic rule is given by
\begin{equation}
\sigma_{e}^{\text{CDW}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}^{\text{CDW}}}{\sum\limits_{\alpha}n_{\alpha}
\nu_{\alpha}},
\label{32}
\end{equation}
\textbf{where $\nu_{\alpha}$ are the new numbers of active electrons per 
atom obtained from the CDW ionization cross sections for different ions 
in H, C, N, O, P , and S targets. The new values are
\begin{equation}
\nu_{\alpha}=\left\{
\begin{array}{ll}
1, & \text{H,} \\
4, & \text{for C, N, O} \\
4.5, & \text{for P, S} \\
\end{array}\right.
\label{eq:scalingCDW}
\end{equation}
}
The \textbf{scaled} cross sections
$\sigma_{e}^{\text{CDW}}$ are plotted in Figure~4b. A much better sharp
band is observed, especially \textbf{for impact energies $E=(0.5-8)MeV/amu$ 
for $Z=1$ and $Z=2$; and $E=(2.5-8)MeV/amu$ for $Z>2$.}
%at high energy, where the theory is expected to work.
\textbf{
The experimental data for ionization of adenine, cytosine, thymine, 
guanine, uracil, pyrimidine and THF by proton impact seems to 
corroborate the new scaling. It will be interesting to cross-check for 
experiments with projectile charge states $Z>1$.}

\textbf{In table \ref{nn} we display the number of active electrons 
obtained using the present scaling for some molecules of interest. These 
numbers are very different from those proposed by Toburen \cite{} and 
used by other authors \cite{Itoh,others}.}

\begin{table}
\begin{center}
\begin{tabular}{llllll}
\hline
 Molecule & number &Molecule         &number & Molecule             &number \\
 H$_2$    & 2  & C$_2$H$_7$N         & 19    & C$_4$H$_5$N$_3$O     & 37    \\
 H$_2$O   & 6  & C$_4$H$_8$O         & 28    & C$_5$H$_6$N$_2$O$_2$ & 42    \\
 NH$_3$   & 7  & C$_4$H$_4$N$_2$     & 28    & C$_5$H$_5$N$_5$      & 45    \\
 CH$_4$   & 8  & C$_6$H$_6$          & 30    & C$_5$H$_5$N$_5$O     & 49    \\
 CH$_5$N  & 13 & C$_4$H$_4$N$_2$O$_2$& 36    & C$_5$H$_{10}$O$_5$P  & 54.5  \\
 \hline
\end{tabular}
\caption{New scaling numbers for SOME molecular targets of biological interest.}
\label{nn}
\end{center}
\end{table}

\textbf{An alternative way of showing this scaling is to plot the ionization cross sections of molecules as a function of the number of active electrons given in table \ref{nn}. This is displayed in the figure \ref{recta} for impact energies 0.5, 1 and 2 MeV. As can be noted the CDW ionization cross sections calculated for different molecules actually show a linear dependence with the number of electrons given in table \ref{nn}. We obtain similar straigth lines even for E=10 MeV. The comparisson with the experimental data available shows very nice agreement from the smallest molecules H$_2$, H$_2$O and CH$_4$, up to the most complex ones like adenine or guanine}.

\begin{figure}[H]
\centering
\includegraphics[width=0.95\textwidth]{fig_recta.eps}
\end{figure}
